### 2020-06-17 manuscript/internship_report.Rmd
add Ehinger et al (2019) results

# Algorithm Development
As part of answering my research questions, I developed an algorithm named gazeHMM to classify gaze data into discrete eye movement events. The following section describes the development of gazeHMM and its final version that has been used to obtain results. Technical details can be found in the Technical Appendix.

## Eye Movement Metrics
Many different metrics can be used to describe gaze data and eye movement events [@Zemblys2018]. Here, the goal is to find those metrics that separate the gaze data samples belonging to different events the best. However, many metrics rely on thresholds or window ranges that have to be set by the user [e.g., the distance between the mean position in a 100 ms window before and after each sample, see @Olsson2007; @Zemblys2018]. This can be problematic because these parameters are often set without theoretical justification and they differ substantially between metrics. Belonging to the most basic metrics, which do not require parameter settings, I used velocity, acceleration, and sample-to-sample angle [similar to relative or change in angle; @Larsson2013] in gazeHMM.  
Theoretically, these three metrics should separate eye movement events clearly. Fixations typically inherit samples with low velocity and acceleration [@Larsson2013]. Due to tremor, the angle between samples should not follow any direction but a random walk [@Duchowski2017]. In contrast, saccade samples usually have a high velocity and acceleration and follow the same direction. PSO samples tend to have moderate velocity and high acceleration since they occur after high velocity saccades [@Larsson2013]. They are specifically distinguished by their change in direction clustered around 180° [@Pekkanen2017]. Lastly, smooth pursuit samples have a moderate velocity but low acceleration (due to the smoothness) and like saccades they follow the same direction [@Larsson2013; @Leigh2015].

@@

```{r, filter-angle, fig.cap="(ref:filter-angle)"}

@@

Instead of filtering before computing the metrics, I decided to combine both in one step. Previous algorithms have used two methods that both filter and compute derivatives of the gaze position (i.e., velocity and acceleration). @Houpt2018 compute the first and second discrete derivative (similar to a Sobel and Laplace filter, respectively) of the gaze position to be used by their algorithm. Similarly, @Nystrom2010 use a Savitzky-Golay (SG) filter @Savitzky1964 to estimate velocity and acceleration from a gaze signal. Figure \@ref(fig:filter-comparison) displays velocity and acceleration signals obtained by both methods. They follow a similar trend, but the SG velocity signal is less noisy and the acceleration signal is consequently lower than the Laplace acceleration signal. Therefore, I chose to implement the SG filter to compute velocity and acceleration signals, because it filters out more noise than the discrete derivatives but still preserves the edges in the signal to distinguish between events. To preserve motor noise in the sample-to-sample angle signal, gazeHMM computes the absolute angle with the backward difference and the first discrete derivative as the forward difference (see Technical Appendix).

@@

```{r, filter-comparison, fig.cap="(ref:filter-comparison)"}

@@

(ref:sim-ranges) Approximate ranges of response variables used to generate parameter values
(ref:sim-ranges-note) Units are °/s (velocity), °/s^2^ (acceleration), and radians (angle). ~ indicates that the distribution has a peak at this value. Velocity ranges are based on event velocities reported in @Larsson2013. Since I would also like to test the algorithm on extreme data distributions, I extended the ranges beyond those found in typical eye movement data.

```{r, sim-ranges}

@@

```{r, sim-parameters-trans}

@@

```{r, sim-parameters-resp}

@@

```{r, plot-acc-1-2, fig.cap="(ref:plot-acc-1-2)"}

@@

```{r, plot-acc-1-3, fig.cap="(ref:plot-acc-1-3)"}

@@

```{r, plot-acc-1-4, fig.cap="(ref:plot-acc-1-4)"}

@@

```{r, plot-acc-2, fig.cap="(ref:plot-acc-2)"}

@@

```{r, plot-acc-4, fig.cap="(ref:plot-acc-4)"}

@@

## Validation
### Starting Values and Model Fitting
I applied gazeHMM on two benchmark data sets to test my hypotheses. For both data sets, I used the same set of starting values (initial state model: $\rho=1/k$; transition model: $a_{i=j}=0.9$ and $a_{i\neq j}=0.1/k$; response model: see Table \@ref(tab:tab-starting-values)) to estimate the parameters of the HMM. In contrast to the simulation study, generating random starting value often led to bad model fits and label switching between states. To improve the fitting of the gamma distributions, velocity and acceleration signals were scaled down by factor 100 (so were the starting values for their gamma distributions). The algorithm was applied separately for every participant and every condition. For the data set by @Andersson2017, the algorithm was successfully fitted for every participant and condition.

@@

To examine whether HMMs are accurately describing eye movements, I applied gazeHMM with one, two, three, four, and five states on two benchmark data sets and compared the Schwarz weights of the HMMs. Schwarz weights are the result of transforming a set of BIC values. They can be interpreted as the probability for a model to have generated the data it was fitted to. For the @Andersson2017 data set, I expected the highest weights for the three-state models in the image condition. For the moving dots and video conditions, I predicted the highest weights for the four-state model. Figure \@ref(fig:plot-schwarz-weights-img) shows the Schwarz weights for different subjects and models in the image condition. For all subjects, the five-state model displays the highest weight, suggesting that it hast most likely generated the data. In the moving dots condition, the five-state model has the highest weights for most subjects, but the one-, three, and four-state models exhibit the highest weights for two subjects each (see Figure \@ref(fig:plot-schwarz-weights-dots)). The video condition has the same pattern as the image condition, as the five-state model consistently has the highest weights (see Figure \@ref(fig:plot-schwarz-weights-vid)). In contrast to my hypothesis, applying different HMMs to the @Andersson2017 data suggests that the five-state model has most likely generated the data. Except for the moving dots condition, there was little variation in the Schwarz weights, indicating large differences in the model likelihoods. The model comparison for the @Ehinger2019 data set yielded similar results (see Appendix). A recent model recovery study showed that the BIC tended to prefer overly complex HMMs when they were misspecified [e.g., the conditional independence assumption was violated; @Pohle2017]. Instead, the integrated completed likelihood (ICL) criterion [@Biernacki2013] performed better in choosing the correct data generating model. Therefore, I computed the weighted ICL criterion for the HMMs fitted to the benchmark data sets. Using the ICL as the model selection criterion yielded very similar results to the BIC (see Appendix). The preference for the five-state model was even more consistent across conditions and subjects.

@@

(1) *There were more eye movement events present in the data than I expected*. Eye movement events can be divided into subevents. For example, fixations consist of drift and tremor movements [@Duchowski2017] and PSOs encompass dynamic, static, and glissadic over- and undershoots [@Larsson2013]. A study on a recently developed HMM algorithm supports this explanation: @Houpt2018 applied the unsupervised BP-AR-HMM algorithm to the @Andersson2017 data set and classified more distinct states than the human coders. Some of the states classified by BP-AR-HMM matched the same event coded by humans. Since the subevents are usually not interesting for users of classification algorithms, the ability of HMMs to classify might limit their ability to describe eye movements here.  
(2) *Model selection criteria are generally not appropriate for comparing HMMs with differnt numbers of states*. This argument has been discussed in the field of ecology [see @Li2017], where studies found that selection criteria preferred models with more states than expected [similar to the result of this study; e.g., @Langrock2015]. @Li2017 explain this bias with the simplicity of the submodels in HMMs: Initial state, transition, and response models for each state are usually relatively simple. When they do not describe the processes in the respective states accurately, the selection criteria compensate for that by preferring a model with more states. Thus, there are not more latent states present in the data, but the submodels of the HMM are misspecified or too simple. Correcting for model misspecifications, led to better model recovery in studies on animal movements [@Langrock2015; @Li2017]. However, @Pohle2017 showed in simulations that the ICL identified the correct model despite several misspecifications. It has to be noted that the study by @Pohle2017 only used data generating models with two states.  
(3) *The submodels of gazeHMM were misspecified*. @Pohle2017 identified two scenarios in which model recovery using the ICL did not give optimal results: Outliers in the data and inadequate distributions in the response models. Both situations could apply to gazeHMM and eye movement data. Outliers occur frequently in eye-tracking data due to measurement error. Choosing adequate response distributions in HMMs is usually difficult and can depend on the individual and task the data is obtained from [@Langrock2015]. Moreover, gazeHMM only estimated intercepts for all parameters and thus, no time-varying covariates were included [cf. @Li2017]. This aspect could indeed oversimplify the complex nature of eye movement data.

@@

## Advantages of gazeHMM
In view of the four proposed goals that gazeHMM should fulfill, I can draw the following conclusions: Even though gazeHMM does require some parameter settings (in the pre- and postprocessing), they have only minor influence on the classification and are merely included to compensate for some drawbacks of the generative model. Their default values should be appriate for most applications. A major advantage of gazeHMM is that it does not require human labeled data as input. Instead, it estimates all parameters and hidden states from the data. Since human coding is quite laborious, difficult to reproduce, and by times inconsistent [as noted earlier, @Hooge2018], this property makes gazeHMM a good alternative to other recently developed algorithms that require human coded input [@Pekkanen2017; @Zemblys2018; @Bellet2019]. This could also explain why the agreement to human coding is lower for gazeHMM than for algorithms that learn from human labeled data. Another advantage of gazeHMM is its ability to classify four eye movement events, namely fixations, saccades, PSOs, and smooth pursuit. Whereas most algorithms only parse fixations and saccades [@Andersson2017], few are able to classify PSOs [e.g., @Zemblys2017], and even less categorize smooth pursuits [e.g., @Pekkanen2017]. However, including smooth pursuits in gazeHMM led to some undesirable classifications on benchmark data, resulting in rapid switching between fixation and smooth pursuit events. Therefore, I recommend using gazeHMM with four events only for exploratory purposes. Without smooth pursuits, I consider gazeHMM's classification as appropriate for application. Lastly, its implementation in R using depmixS4 [@Visser2010] should make gazeHMM a tool that is easy to use and customize for individual needs.

## Future Directions
Despite its advantages, there are several aspects in which gazeHMM can be improved: First, a multivariate distribution could be used to account for the correlation between velocity and acceleration signals [for an example, see @Balakrishnan2009]. Potential problem of this approach might be choosing the right distribution and convergence issues (due to a large number of parameters). Another option to model the correlation could be to include one of the response variables as a covariate of the other.  
Second, instead of the gamma being the generic (and potentially inappropriate) response distribution, a non-parametric approach could be used: @Langrock2015 use a linear combination of standardized B-splines to approximate response densities, which led to HMMs with less states being preferred. This approach could potentially combat the problem of unexpectedly high-state HMMs being preferred for eye movement data.  
Third, including covariates of the transition probabilities and response parameters could improve the fit of gazeHMM on eye movement data. As pointed out earlier, just estimating intercepts of parameters could be too simple to model the complexity of eye movements. As candidate covariates I suggest periodic functions of time [@Li2017] which could, for instance, capture the specific pattern of saccades.  
Fourth, to avoid rapid switching between fixations and smooth pursuits as well as unreasonable short saccades, gazeHMM could explicitly model the duration of events. This can be achieved by setting the diagonal transition probabilities to zero and assign a distribution of state durations to each state [@Bishop2015]. Consequently, the duration distributions of fixations and smooth pursuits could differ from saccades and PSOs. This extension of the HMM is also called hidden semi-Markov model and has been successfully used by @Mihali2017 to classify microsaccades. A drawback of the extension would be higher computational costs.  
Lastly, allowing constrained parameters in the HMM could replace some of the postprocessing steps in gazeHMM. This could potentially be achieved by using different response distributions or parameter optimization methods. Moreover, switching from the maximum likelihood to the Markov chain Monte Carlo (Bayesian) framework could help avoiding convergence problems with constrained parameters, but would also open new research questions about suitable priors for HMM parameters in the eye movement domain.

@@

*Acknowledgment*. I would like to thank my supervisors Ingmar Visser and Simon Kucharsky for their helpful advice on the development and implementation of the study.

>>>>

# Algorithm Development
As part of answering my research questions, I developed an algorithm named gazeHMM to classify gaze data into discrete eye movement events. The following section describes the development of gazeHMM and its final version that has been used to obtain results. Technical details can be found in Appendix B.

## Eye Movement Metrics
Many different metrics can be used to describe gaze data and eye movement events [@Zemblys2018]. Here, the goal is to find those metrics that separate the gaze data samples belonging to different events the best. However, many metrics rely on thresholds or window ranges that have to be set by the user [e.g., the distance between the mean position in a 100 ms window before and after each sample, see @Olsson2007; @Zemblys2018]. This can be problematic because these parameters are often set without theoretical justification and they differ substantially between metrics. Belonging to the most basic metrics, which do not require parameter settings, I used velocity, acceleration, and sample-to-sample angle [similar to relative or change in angle; @Larsson2013] in gazeHMM.  
Theoretically, these three metrics should separate eye movement events clearly. Fixations typically inherit samples with low velocity and acceleration [@Larsson2013]. Due to tremor, the angle between samples should not follow any direction but a random walk [@Duchowski2017]. In contrast, saccade samples usually have a high velocity and acceleration and follow the same direction. PSO samples tend to have moderate velocity and high acceleration since they occur after high velocity saccades [@Larsson2013]. They are specifically distinguished by their change in direction clustered around 180Ã‚Â° [@Pekkanen2017]. Lastly, smooth pursuit samples have a moderate velocity but low acceleration (due to the smoothness) and like saccades they follow the same direction [@Larsson2013; @Leigh2015].

@@

```{r filter-angle, fig.cap="(ref:filter-angle)"}

@@

Instead of filtering before computing the metrics, I decided to combine both in one step. Previous algorithms have used two methods that both filter and compute derivatives of the gaze position (i.e., velocity and acceleration). @Houpt2018 compute the first and second discrete derivative (similar to a Sobel and Laplace filter, respectively) of the gaze position to be used by their algorithm. Similarly, @Nystrom2010 use a Savitzky-Golay (SG) filter @Savitzky1964 to estimate velocity and acceleration from a gaze signal. Figure \@ref(fig:filter-comparison) displays velocity and acceleration signals obtained by both methods. They follow a similar trend, but the SG velocity signal is less noisy and the acceleration signal is consequently lower than the Laplace acceleration signal. Therefore, I chose to implement the SG filter to compute velocity and acceleration signals, because it filters out more noise than the discrete derivatives but still preserves the edges in the signal to distinguish between events. To preserve motor noise in the sample-to-sample angle signal, gazeHMM computes the absolute angle with the backward difference and the first discrete derivative as the forward difference (see Appendix B).

@@

```{r filter-comparison, fig.cap="(ref:filter-comparison)"}

@@

(ref:sim-ranges) Approximate ranges of response variables used to generate parameter values
(ref:sim-ranges-note) Units are Ã‚Â°/s (velocity), Ã‚Â°/s^2^ (acceleration), and radians (angle). ~ indicates that the distribution has a peak at this value. Velocity ranges are based on event velocities reported in @Larsson2013. Since I would also like to test the algorithm on extreme data distributions, I extended the ranges beyond those found in typical eye movement data.

```{r sim-ranges}

@@

```{r sim-parameters-trans}

@@

```{r sim-parameters-resp}

@@

```{r plot-acc-1-2, fig.cap="(ref:plot-acc-1-2)"}

@@

```{r plot-acc-1-3, fig.cap="(ref:plot-acc-1-3)"}

@@

```{r plot-acc-1-4, fig.cap="(ref:plot-acc-1-4)"}

@@

```{r plot-acc-2, fig.cap="(ref:plot-acc-2)"}

@@

```{r plot-acc-4, fig.cap="(ref:plot-acc-4)"}

@@

## Validation
### Starting Values and Model Fitting
I applied gazeHMM on two benchmark data sets to test my hypotheses. Note that, for the @Ehinger2019 data set, I only fitted models to the Eyelink data to reduce the complexity of the analyses. For both data sets, I used the same set of starting values (initial state model: $\rho=1/k$; transition model: $a_{i=j}=0.9$ and $a_{i\neq j}=0.1/k$; response model: see Table \@ref(tab:tab-starting-values)) to estimate the parameters of the HMM. In contrast to the simulation study, generating random starting value often led to bad model fits and label switching between states. To improve the fitting of the gamma distributions, velocity and acceleration signals were scaled down by factor 100 (so were the starting values for their gamma distributions). The algorithm was applied separately for every participant and every condition, task, or block. For the data set by @Andersson2017, the algorithm was successfully fitted for every participant and condition.

@@

To examine whether HMMs are accurately describing eye movements, I applied gazeHMM with one, two, three, four, and five states on two benchmark data sets and compared the Schwarz weights of the HMMs. Schwarz weights are the result of transforming a set of BIC values. They can be interpreted as the probability for a model to have generated the data it was fitted to. For the @Andersson2017 data set, I expected the highest weights for the three-state models in the image condition. For the moving dots and video conditions, I predicted the highest weights for the four-state model. Figure \@ref(fig:plot-schwarz-weights-img) shows the Schwarz weights for different subjects and models in the image condition. For all subjects, the five-state model displays the highest weight, suggesting that it hast most likely generated the data. In the moving dots condition, the five-state model has the highest weights for most subjects, but the one-, three, and four-state models exhibit the highest weights for two subjects each (see Figure \@ref(fig:plot-schwarz-weights-dots)). The video condition has the same pattern as the image condition, as the five-state model consistently has the highest weights (see Figure \@ref(fig:plot-schwarz-weights-vid)). In contrast to my hypothesis, applying different HMMs to the @Andersson2017 data suggests that the five-state model has most likely generated the data. Except for the moving dots condition, there was little variation in the Schwarz weights, indicating large differences in the model likelihoods. The model comparison for the @Ehinger2019 data set yielded similar results (see Figures \@ref(fig:plot-schwarz-weights-task4) and \@ref(fig:plot-schwarz-weights-task5)). A recent model recovery study showed that the BIC tended to prefer overly complex HMMs when they were misspecified [e.g., the conditional independence assumption was violated; @Pohle2017]. Instead, the integrated completed likelihood (ICL) criterion [@Biernacki2000] performed better in choosing the correct data generating model. Therefore, I computed the weighted ICL criterion for the HMMs fitted to the @Andersson2017 data set. Using the ICL as the model selection criterion yielded very similar results to the BIC (see Figures \@ref(fig:plot-icl-weights-img), \@ref(fig:plot-icl-weights-dots), and \@ref(fig:plot-icl-weights-vid)). The preference for the five-state model was even more consistent across conditions and subjects.

@@

*There were more eye movement events present in the data than I expected*. Eye movement events can be divided into subevents. For example, fixations consist of drift and tremor movements [@Duchowski2017] and PSOs encompass dynamic, static, and glissadic over- and undershoots [@Larsson2013]. A study on a recently developed HMM algorithm supports this explanation: @Houpt2018 applied the unsupervised BP-AR-HMM algorithm to the @Andersson2017 data set and classified more distinct states than the human coders. Some of the states classified by BP-AR-HMM matched the same event coded by humans. Since the subevents are usually not interesting for users of classification algorithms, the ability of HMMs to classify might limit their ability to describe eye movements here.  
*Model selection criteria are generally not appropriate for comparing HMMs with differnt numbers of states*. This argument has been discussed in the field of ecology [see @Li2017], where studies found that selection criteria preferred models with more states than expected [similar to the result of this study; e.g., @Langrock2015]. @Li2017 explain this bias with the simplicity of the submodels in HMMs: Initial state, transition, and response models for each state are usually relatively simple. When they do not describe the processes in the respective states accurately, the selection criteria compensate for that by preferring a model with more states. Thus, there are not more latent states present in the data, but the submodels of the HMM are misspecified or too simple. Correcting for model misspecifications, led to better model recovery in studies on animal movements [@Langrock2015; @Li2017]. However, @Pohle2017 showed in simulations that the ICL identified the correct model despite several misspecifications. It has to be noted that the study by @Pohle2017 only used data generating models with two states.  
*The submodels of gazeHMM were misspecified*. @Pohle2017 identified two scenarios in which model recovery using the ICL did not give optimal results: Outliers in the data and inadequate distributions in the response models. Both situations could apply to gazeHMM and eye movement data. Outliers occur frequently in eye-tracking data due to measurement error. Choosing adequate response distributions in HMMs is usually difficult and can depend on the individual and task the data is obtained from [@Langrock2015]. Moreover, gazeHMM only estimated intercepts for all parameters and thus, no time-varying covariates were included [cf. @Li2017]. This aspect could indeed oversimplify the complex nature of eye movement data.

@@

## Advantages of gazeHMM
In view of the four proposed goals that gazeHMM should fulfill, I can draw the following conclusions: Even though gazeHMM does require some parameter settings (in the pre- and postprocessing), they have only minor influence on the classification and are merely included to compensate for some drawbacks of the generative model. Their default values should be appriate for most applications. A major advantage of gazeHMM is that it does not require human labeled data as input. Instead, it estimates all parameters and hidden states from the data. Since human coding is quite laborious, difficult to reproduce, and by times inconsistent [as noted earlier, @Hooge2018], this property makes gazeHMM a good alternative to other recently developed algorithms that require human coded input [@Pekkanen2017; @Zemblys2018; @Bellet2019]. This could also explain why the agreement to human coding is lower for gazeHMM than for algorithms that learn from human labeled data. Another advantage of gazeHMM is its ability to classify four eye movement events, namely fixations, saccades, PSOs, and smooth pursuit. Whereas most algorithms only parse fixations and saccades [@Andersson2017], few are able to classify PSOs [e.g., @Zemblys2018], and even less categorize smooth pursuits [e.g., @Pekkanen2017]. However, including smooth pursuits in gazeHMM led to some undesirable classifications on benchmark data, resulting in rapid switching between fixation and smooth pursuit events. Therefore, I recommend using gazeHMM with four events only for exploratory purposes. Without smooth pursuits, I consider gazeHMM's classification as appropriate for application. Lastly, its implementation in R using depmixS4 [@Visser2010] should make gazeHMM a tool that is easy to use and customize for individual needs.

## Future Directions
Despite its advantages, there are several aspects in which gazeHMM can be improved: First, a multivariate distribution could be used to account for the correlation between velocity and acceleration signals [for an example, see @Balakrishnan2009]. Potential problem of this approach might be choosing the right distribution and convergence issues (due to a large number of parameters). Another option to model the correlation could be to include one of the response variables as a covariate of the other.  
Second, instead of the gamma being the generic (and potentially inappropriate) response distribution, a non-parametric approach could be used: @Langrock2015 use a linear combination of standardized B-splines to approximate response densities, which led to HMMs with less states being preferred. This approach could potentially combat the problem of unexpectedly high-state HMMs being preferred for eye movement data.  
Third, including covariates of the transition probabilities and response parameters could improve the fit of gazeHMM on eye movement data. As pointed out earlier, just estimating intercepts of parameters could be too simple to model the complexity of eye movements. As candidate covariates I suggest periodic functions of time [@Li2017] which could, for instance, capture the specific pattern of saccades.  
Fourth, to avoid rapid switching between fixations and smooth pursuits as well as unreasonable short saccades, gazeHMM could explicitly model the duration of events. This can be achieved by setting the diagonal transition probabilities to zero and assign a distribution of state durations to each state [@Bishop2006]. Consequently, the duration distributions of fixations and smooth pursuits could differ from saccades and PSOs. This extension of the HMM is also called hidden semi-Markov model and has been successfully used by @Mihali2017 to classify microsaccades. A drawback of the extension would be higher computational costs.  
Lastly, allowing constrained parameters in the HMM could replace some of the postprocessing steps in gazeHMM. This could potentially be achieved by using different response distributions or parameter optimization methods. Moreover, switching from the maximum likelihood to the Markov chain Monte Carlo (Bayesian) framework could help avoiding convergence problems with constrained parameters, but would also open new research questions about suitable priors for HMM parameters in the eye movement domain.

@@

### 2020-06-17 manuscript/report_appendix.Rmd
add Ehinger et al (2019) results

(ref:plot-sim-exp) Simulated data from the HMM with the default parameters of the simulation study. Displayed are acceleration (in deg/s$^2$, multiplied by 10000), sample-to-sample (relative) angle (in rad), and velocity (in deg/s, multiplied by 10) over time (in s). Acceleration and velocity were multiplied back to their natural scale. Note that PSOs (State 3, blue dots) do not only occur after saccades because the HMM parameters were unconstrained.

```{r plot-sim-exp, results='hold', fig.cap="(ref:plot-sim-exp)"}

source(here("simulation/model_simulation.R"))

@@

@@

(ref:plot-icl-weights-img) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the image condition of the @Andersson2017 data set. Higher weights indicate better model fit.

```{r plot-icl-weights-img, results='hold', fig.cap="(ref:plot-icl-weights-img)"}

@@

(ref:plot-icl-weights-dots) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the moving dots condition of the @Andersson2017 data set. Higher weights indicate better model fit.

```{r plot-icl-weights-dots, results='hold', fig.cap="(ref:plot-icl-weights-dots)"}

@@

(ref:plot-icl-weights-vid) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the video condition of the @Andersson2017 data set. Higher weights indicate better model fit.

```{r plot-icl-weights-vid, results='hold', fig.cap="(ref:plot-icl-weights-vid)"}

@@

(ref:plot-example-classification) Example data from @Andersson2017 displayed as x-, and y-coordinates, velocity (in deg/s), acceleration (in deg/s$^2$), and sample-to-sample (relative) angle (in rad). Colors indicate classification by gazeHMM and symbols by human coder MN. Note that the human coder classified all fixations from gazeHMM as smooth pursuits and delayed the separation between the saccade and PSOs.

```{r plot-example-classification, results='hold', fig.cap="(ref:plot-example-classification)", fig.height=8}

@@

(ref:tab-example-trans) Transition probabability estimates for example data from @Andersson2017
(ref:tab-example-trans-note) Rows indicate from which state and columns into which state the model transitions.

```{r tab-example-trans, results='hold'}

@@

(ref:tab-example-resp) Response parameter estimates for example data from @Andersson2017
(ref:tab-example-resp-note) Con = concentration parameter of the von-Mises distribution.

```{r tab-example-resp, results='hold'}

>>>>

# Additional Figures and Tables

(ref:plot-sim-ex) Simulated data from the HMM with the default parameters of the simulation study. Displayed are acceleration (in deg/s$^2$, multiplied by 10000), sample-to-sample (relative) angle (in rad), and velocity (in deg/s, multiplied by 10) over time (in s). Acceleration and velocity were multiplied back to their natural scale. Note that PSOs (State 3, blue dots) do not only occur after saccades because the HMM parameters were unconstrained.

```{r plot-sim-ex, fig.cap="(ref:plot-sim-ex)"}

source(here::here("simulation/model_simulation.R"))

@@

```{r calculate Schwarz weights for Ehinger et al., include=FALSE}

load(here("validation/Ehinger2019_fitted.Rdata"))

E2019.bic <- lapply(E2019.fit, function(subj) {
  out <- lapply(subj, function(task) {
    out <- lapply(task, function(block) {
      out <- lapply(1:length(block), function(mod) {

          bic <- try(-2*block[[mod]][[1]] + 6*log(block[[mod]][[2]]))

          bic <- try(BIC(block[[mod]]$model))

        return(ifelse(is.numeric(bic), bic, NA))
      })

    names(df) <- paste("model_", 1:ncol(df), sep = "")

    #blockm <- apply(df, 2, median, na.rm = T)

  return(out)
})

# Create Schwarz weight plots

bic.plot <- list()

bic.plot[[4]] <- reduce(reduce(E2019.bic, rbind)[,1], rbind) %>%
  mutate(Subject = rep(1:15, each = 6),
         Block = rep(1:6, 15)) %>%
  pivot_longer(contains("model"), names_to = "model", values_to = "weight") %>%
  mutate_at(c("Subject", "Block", "model"), as.factor) %>%
  ggplot(aes(x = model, y = Subject)) + geom_tile(aes(fill = weight)) +
  facet_wrap(vars(Block), ncol = 2, labeller = label_both) +
  scale_x_discrete(name = "Number of states", labels = as.character(1:5)) +
  scale_y_discrete(name = "Subject") +
  scale_fill_distiller(name = "Schwarz\nweight", breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1), 
                       palette = "Spectral")

bic.plot[[5]] <- reduce(reduce(E2019.bic, rbind)[,2], rbind) %>%
  mutate(Subject = rep(1:15, each = 6),
         Block = rep(1:6, 15)) %>%
  pivot_longer(contains("model"), names_to = "model", values_to = "weight") %>%
  mutate_at(c("Subject", "Block", "model"), as.factor) %>%
  ggplot(aes(x = model, y = Subject)) + geom_tile(aes(fill = weight)) +
  facet_wrap(vars(Block), ncol = 2, labeller = label_both) +
  scale_x_discrete(name = "Number of states", labels = as.character(1:5)) +
  scale_y_discrete(name = "Subject") +
  scale_fill_distiller(name = "Schwarz\nweight", breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1), 
                       palette = "Spectral")

(ref:plot-schwarz-weights-task4) Schwarz weights displayed for each subject and HMMs with different numbers of states. Models were applied to task 4 of the Ehinger et al. (2019) data set. Higher weights indicate better model fit. Grey tiles indicate erroneous model fits.

```{r plot-schwarz-weights-task4, fig.cap="(ref:plot-schwarz-weights-task4)", fig.height=8}

print(bic.plot[[4]])

(ref:plot-schwarz-weights-task5) Schwarz weights displayed for each subject and HMMs with different numbers of states. Models were applied to task 5 of the Ehinger et al. (2019) data set. Higher weights indicate better model fit. Grey tiles indicate erroneous model fits.

```{r plot-schwarz-weights-task5, fig.cap="(ref:plot-schwarz-weights-task5)", fig.height=8}

print(bic.plot[[5]])

@@

(ref:plot-icl-weights-img) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the image condition of the Andersson et al. (2017) data set. Higher weights indicate better model fit.

```{r plot-icl-weights-img, fig.cap="(ref:plot-icl-weights-img)"}

@@

(ref:plot-icl-weights-dots) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the moving dots condition of the Andersson et al. (2017) data set. Higher weights indicate better model fit.

```{r plot-icl-weights-dots, fig.cap="(ref:plot-icl-weights-dots)"}

@@

(ref:plot-icl-weights-vid) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the video condition of the Andersson et al. (2017) data set. Higher weights indicate better model fit.

```{r plot-icl-weights-vid,  fig.cap="(ref:plot-icl-weights-vid)"}

@@

(ref:plot-example-classification) Example data from Andersson et al. (2017) displayed as x-, and y-coordinates, velocity (in deg/s), acceleration (in deg/s), and sample-to-sample (relative) angle (in rad). Colors indicate classification by gazeHMM and symbols by human coder MN. Note that the human coder classified all fixations from gazeHMM as smooth pursuits and delayed the separation between the saccade and PSOs.

```{r plot-example-classification, fig.cap="(ref:plot-example-classification)", fig.height=7.5}

@@

(ref:tab-example-trans) Transition probabability estimates for example data from Andersson et al. (2017)
(ref:tab-example-trans-note) Rows indicate from which state and columns into which state the model transitions.

```{r tab-example-trans}

@@

(ref:tab-example-resp) Response parameter estimates for example data from Andersson et al. (2017)
(ref:tab-example-resp-note) Con = concentration parameter of the von-Mises distribution.

```{r tab-example-resp}
